---
title: "Schedule for Statistical inference, HT19"
author: "Martin Sköld"
date: 'Last update on `r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, echo = FALSE, message=FALSE}
library(jsonlite)
library(tidyverse)

# Byt https mot aktuellt schemalänk nedan
datum <- unique(fromJSON("https://cloud.timeedit.net/su/web/stud1/ri177455X45Z06Q5Z96g6Y00y5016Y39Q07gQY6Q53767.json", flatten = TRUE)$reservations) %>% 
    filter(str_detect(columns, "Föreläsning")) %>% 
    mutate(veckodag = weekdays(as.Date(startdate))) %>% 
    select(veckodag, startdate) %>% 
    distinct()
```


***

**`r datum[1,]`**

*Introduction. Statical models and likelihood.*  
Literature: Chapter 1, Chapter 2-2.2.  
Recommended exercises:  2.6: 1, 2, 4. Ö: 4, 11, 12.  
Data for 1d: `x <- c(1.5, 0.25, 3.75, 3.0, 2.5)`  
Data for 2c: `x <- c(-0.560, -0.510, 1.304, 0.722, 0.490, 1.960, 1.441)`  


***

**`r datum[2,]`**

*Score and information, Numerical methods, sufficiency.*  
Literature: Chapter 2.2-2.5. (ej 2.3.2)  
Recommended exercises: 2.6: 5 (in 5c the score-equation $S=0$ has two solutions, but only one in the parameter space). Ö: 1, 3.  

***

**`r datum[3,]`**

*Sufficiency. Topics in frequentist statistics.*  
Literature: Chapter 2.5, 3.1-2.  
Recommended exercises: 2.6: 9, 6, 7, 8. Ö: 2a.  

***

**`r datum[4,]`**

*Topics in frequentist statistics, asymptotics. the Bootstrap.*  
Literature: Chapter 3-3.2.5  
Recommended exercises: 3.4: 2, 3, 5, 7, 8. Ö: 6.  

***

**`r datum[5,]`**

*Tests and P-values.*  
Literature: Chapter 3.3.  
Recommended exercises: Ö: 7-8.  

***

**`r datum[6,]`**

*Frequentist likelihood theory: Methods based on Score.*  
Literature: Chapter 4-4.2.  
Recommended exercises: 4.7: 2, 5, 6a,b (also derive a 95% score-interval by plotting the score-statistic as a function of $\pi$ and visually approximate the limits, data in 6f)


***

**`r datum[7,]`**

*Frequentist likelihood theory: Wald and Likelihood ratios.*  
Literature: Chapter 4.2.  
Recommended exercises: 4.7: 3, 6c-g, 7.  

***

**`r datum[8,]`**

*Frequentist likelihood theory.*  
Literature:  Chapter 4.4, 4.6.  
Recommended exercises: [Tentamen 20160224](http://kurser.math.su.se/pluginfile.php/1460/mod_folder/content/0/20160224.pdf?forcedownload=1), 1bc, 3ab. [Tentamen 20160113](http://kurser.math.su.se/pluginfile.php/1460/mod_folder/content/0/20160113.pdf?forcedownload=1), 1bc, 3ab. [Tentamen 20150116](http://kurser.math.su.se/pluginfile.php/1460/mod_folder/content/0/20150116.pdf?forcedownload=1), 3cd.  

***

**`r datum[9,]`**

*Likelihood theory for vector valued parameters.*  
Literature: Chapter 5.1-5.4  
Recommended exercises: 5.7:1, 10, 12.  

***

**`r datum[10,]`**

*Likelihood theory for vector valued parameters, Bayesian statistics.*  
Literature: 5.4-5.5, 6.1-6.2.  
Recommended exercises: 5.7: 13, 15 a-d (when $K=2$).  
Data for 5.7:13:  
`x <- c(578,835,759,1940,1769,1030,2343,484,1803,1060,1096,3365,3470,256,443,2335,438,150,963,1584,2537,2174,1114,1023,1217,596,533,1781,2046,1367,48,2267,729,2052,1975,2763,1744,302,352,38,1335,2531,3039,1693,3337,1024,3382,1237,395,1275,1102,1665,2843,456,457,245,2593,522,1881,1189,1219,1825,2431,2570,586,1560,1927,1202,2429,1708,785,168,2046,963,535,2010,1159,2425,1528,1056,2212,1686,19,1816,2603,2183,1184,1919,9,341)`  
`y <- c(3299,962,1436,769,335,191,4253,3912,193,251,743,3703,2500,2929,2529,630,1942,4108,1499,754,647,3906,2470,974,2857,3024,38,1268,421,688,96,3456,464,1554,1626,2024,1998,2063,574,797,3056,636,1500,3414,3247,616,2101,1555,1552,1649,1247,1219,947,201,799,144,425,2415,582,828,1292,1975,1260,932,2522,2744,9,2387,1522,2338,2384,2426,1408,804,287,379,500,167,651,586,2058,2254,2436,2495,1982,2209,1113,904,177,8,498,207,2114,2148)`  
`delta <- c(1,1,1,1,1,1,1,1,0,0,0,1,0,0,1,0,0,1,0,0,0,1,0,0,0,0,0,0,0,0,0,1,0,0,1,1,1,1,1,0,1,0,0,1,1,0,1,1,0,1,0,0,0,0,1,0,0,1,1,1,1,0,0,1,0,1,0,1,0,1,1,0,0,0,1,1,0,0,1,0,1,1,1,1,1,0,1,1,0,1,1,0,1,0)`  
`gamma <- c(0,0,0,0,1,1,0,0,0,0,0,1,1,0,0,1,1,0,0,0,0,1,1,0,0,1,0,0,1,1,0,0,1,1,0,1,0,1,1,0,1,0,1,0,1,0,0,1,0,1,1,0,0,0,0,0,1,0,0,0,1,0,1,1,1,0,0,0,1,1,1,1,1,1,0,1,1,1,0,0,0,1,0,1,0,1,1,1,0,0)`  


***

**`r datum[11,]`**

*Bayesian statistics, choice of  prior and vector valued parameters. MCMC.*  
Literature: Chapter 6.3, 6.5, 8.4.  
Recommended exercises: 6.8: 1, 3 , 4, 8, 9. [Tentamen 20160224](http://kurser.math.su.se/pluginfile.php/1460/mod_folder/content/0/20160224.pdf?forcedownload=1), 2b. [Tentamen 20160113](http://kurser.math.su.se/pluginfile.php/1460/mod_folder/content/0/20160113.pdf?forcedownload=1), 2b. [Tentamen 20150116](http://kurser.math.su.se/pluginfile.php/1460/mod_folder/content/0/20150116.pdf?forcedownload=1), 2ab.  
Data for 6.8:3: `x <- c(183, 173, 181, 170, 176, 180, 187, 176, 171, 190, 184, 173, 176, 179, 181, 186)`.  

***

**`r datum[12,]`**

*Model choice*  
Literature: Chapter 7-7.1.  
Recommended exercises: Ö: 10. 7.3: 2.  

***

**`r datum[13,]`**

*Regularisation, shrinkage and model evaluation.*  
Literature:  ISL Chapter 2.2-2.2.2, 6.2
Recommended exercises:  ISL 2.4:5 and 6.8:2a-b, 3, 4.

***

**`r datum[14,]`**

*Regularisation, shrinkage and model evaluation.*  
Literature:  ISL Chapter 2.2-2.2.2, 6.2
Recommended exercises:  ISL 2.4:5 and 6.8:2a-b, 3, 4.

***

**`r datum[15,]`**

*Repetition.*


